---
output:
  pdf_document:
    keep_tex: yes
---
The Effect of Transmission on Fuel Economy
==========================================
## Executive Summary
The purpose of this analysis is to quantify the relationship between
vehicle transmission and fuel economy, measured in miles per gallon
(mpg). The data we use is taken from the `mtcars` dataset included in
the R programming language . This data is taken from the 1974 Motor
Trend magazine, and details 11 aspects of motor vehicle design and
performance. We find that on average, cars with manual transmissions
are more fuel efficient than those with automatic transmissions by 2.08
miles per gallon. This effect, however, is not statistically significant after
adjusting for variations in weight and horsepower.

## Analysis
This analysis requires the `ggplot2` and `GGally` graphics packages, and the `mtcars` dataset. In Figure 1, we create a matrix plot of the columns of mtcars to see
what relationships there are between the variables. Figure 2 is a boxplot comparing fuel economy between automatic and manual transmission cars.

```{r echo=FALSE}
library(ggplot2)
library(GGally)
data(mtcars)
```

We find many relationships between various columms. This can be problematic for
linear regression; omitting interaction terms can make models less
accurate, but including these terms makes models much harder to
interpret. We will fit a linear model in a stepwise fashion without interaction
terms, starting with transmission type. To limit the effect of the
corelation between variables, we will limit our model to 3 predictors.

First, we create the one-predictor model based on transmission type.

```{r}
mpg_am <- lm(mpg ~ am, data=mtcars)
summary(mpg_am)[c("coefficients","r.squared")]
```

As expected, our model does not do a great job of explaining the
variance in the data set; the R squared is only 0.338.

The next predictor  we add is the one which improves the R squared statistic the most; to find this predictor, we caclulate the linear regression for each remaining column and compare the R squared.

```{r}
variablesLeft <- setdiff(names(mtcars), c("mpg", "am"))
models1 <- lapply(variablesLeft, FUN=function(x) {
    form <- paste0("mpg ~ am + ", x)
    lm(form, data=mtcars)
})
results <- cbind(column=variablesLeft, 
                 adj.r.squared=sapply(models1, 
                                      FUN=function(x) summary(x)$adj.r.squared))
results[order(results[,2], decreasing=TRUE)[1],]
summary(models1[[3]])[c("coefficients","r.squared")]
```

We find that the model evaluating horsepower has the highest R squared
statistic. We repeat the process to find our last predictor to add
to the model.

```{r}
variablesLeft <- setdiff(variablesLeft, "hp")
models2 <- lapply(variablesLeft, FUN=function(x) {
    form <- paste0("mpg ~ am + hp + ", x)
    lm(form, data=mtcars)
})

results <- cbind(column=variablesLeft, 
                 adj.r.squared=sapply(models2, 
                                      FUN=function(x) summary(x)$adj.r.squared))
results[order(results[,2], decreasing=TRUE)[1],]
summary(models2[[4]])[c("coefficients","r.squared")]
 
``` 

We find that weight explains the most variance in the rest of the
data.  Holding these two covariates constant, we find that the
difference between manual and automatic transmission fuel economy is
2.08, but is not significant. Looking at the Residuals plot in Figure 4, we see that there is a distinct U shape in the plot; this is not unexpected, as we intentionally disregarded the effect of interactions between the variables and limited the number of predictors in the model. Incorporating interaction terms would likely improve the fit of the model and reduce the pattern in this residuals plot.

The major weakness of this analysis is that by ignoring interaction
terms, we sacrifice power in order to make the
interpretation of the model easier. As such, our conclusion of a lack
of relation between mpg and transmission type must be taken
cautiously. A study with a larger sample size that considers the interaction between the variables may yield better
results.

## Appendix: Figures
```{r Figures, echo=FALSE, cache=TRUE}
# Figure 1: Matrix plot
ggpairs(mtcars, 
        title="Figure 1: Matrix plot of mtcars", 
        upper= list(continuous="smooth", combo="box", discrete="ratio"), 
        lower=list(continuous="smooth", combo="box", discrete = "ratio"), 
        axisLabels="none", 
        params=c(alpha=0.3))
# Figure 2: Box plot
ggplot(mtcars, aes(x=factor(am, labels=c("auto", "manual")), y=mpg)) + 
  geom_boxplot() +
  labs(x="Transmission", 
       y="Fuel Economy (mpg)", 
       title="Figure 2: Fuel Economy vs. Transmission")
# Figure 3: Fuel Economy vs. Horsepower
ggplot(mtcars, aes(x=hp, y=mpg, color=as.factor(am))) +
  geom_point(aes(size=wt)) +
  labs(x = "Horsepower",
       y = "Fuel Economy (mpg)",
       title = "Figure 3: Fuel Economy vs. Horsepower")
# Figure 4: Residuals plot of final model
mod2 <- fortify(models2[[4]])
ggplot(mod2, aes(.fitted, .resid)) + 
  geom_hline(y_intercept = 0) + 
  geom_point() + 
  geom_smooth(se=FALSE, method="loess") +
  labs(x = "Fitted values",
       y = "Residuals",
       title = "Figure 4: Residuals versus Fitted Values")

``` 